Ripple Consensus Review

### 2015-04-24 paper.tex
consensus discussion and diagrams

Ripple has diverged significantly from the original concept of a decentralized
network representing money explicitly as debt relationship between
parties.\ref{btcmag-introducing-ripple} While the concept of a debt
relationship still exists in the form of trust lines between participants, the
bulk of the codebase and developer documentation now focuses\footnote{For
    instance the ``Tutorials'' section of the Ripple website only explains how
    to create transactions that modify the ledger; there is almost no information
available on how to actually use the trustlines feature.} on the use and
maintenance of a global ledger of transactions and account balances.
Additionally a native currency, XRP, has been added, which is used for
anti-spam transaction fees\footnote{A small amount of XRP is irrovocably
destroyed for every transaction.}, as well as to serve as a universal currency.

@@

Authentication is performed with a simple signature scheme; a scripting system
is not available, nor is multisignature support. To implement the latter a
highly complex single-purpose scheme\ref{ripple-wiki-multisign} has been
proposed.

@@

\subsection{Consensus}

Claims from ripple: consensus protocol will maintain correctness so long as $f
\le (n-1)/5$ where $f$ is the number of Byzantian failures. However default UNL
list is just $5$ entries, thus $f \le (5-1)/5 = 0.8$ - no failures can be
tolerated.

@@

can talk about how cost to do a simulation attack is PoW $/targets; talk about slasher rule

>>>>

Ripple has diverged significantly from the original concept of a decentralized
network representing money explicitly as debt relationship between
parties.\cite{btcmag-introducing-ripple} While the concept of a debt
relationship still exists in the form of trust lines between participants, the
bulk of the codebase and developer documentation now focuses\footnote{For
    instance the ``Tutorials'' section of the Ripple website only explains how
    to create transactions that modify the ledger; there is almost no information
available on how to actually use the trustlines feature.} on the use and
maintenance of a global ledger of transactions and account balances.
Additionally a native currency, XRP, has been added, which is used for
anti-spam transaction fees\footnote{A small amount of XRP is irrovocably
destroyed for every transaction.}, as well as to serve as a universal currency.

@@

Authentication is performed with a simple signature scheme; a scripting system
is not available, nor is multisignature support. To implement the latter a
highly complex single-purpose scheme\cite{ripple-wiki-multisign} has been
proposed.

@@

\subsection{Unique Node List}

The actual process of determining what version of the ledger is correct starts
with the Unique Node Listi, (UNL) a list of public keys meant to be associated
with active nodes the node operator believes are
``unique''.\cite{ripple-wiki-unl} Ripple Labs suggests that UNL's ``should have
100+ nodes on them.''\cite{ripple-wiki-unl} and provides a ``starter'' UNL at
\url{https://ripple.com/ripple.txt} with $5$ to $8$ nodes.\footnote{Exactly how this UNL is
generated is unknown; the author downloaded it on multiple occasions getting
between $5$ to $8$ nodes each time, mostly the same.}

Through the Ripple Protocol Consensus Algorithm\cite{ripple-consensus-paper}
nodes on the UNL vote to determine the contents of the Ripple ledger. While the
actual protocol contains a number of rounds of proposals and voting the end
result can be described as basically a supermajority vote: a transaction is
only approved if $80\%$ of the UNL of a server agrees with
it.\cite[3.2]{ripple-consensus-paper} Put another way, if $20\%$ of the UNL
choose to reject a transaction, it will not be included in the ledger.

Different nodes may chose to use different UNLs; if the UNLs do not
sufficiently overlap global consensus is not guaranteed because different UNL
``cliques'' can come to consensus independently of each other. For instance
Figure \ref{fig:unl-clique-disjoint} shows a simple example with two UNL
cliques, red and blue, that have forked because of insufficient connectivity,
resulting in the forked blockchain shown in Figure \ref{fig:unl-clique-disjoint-blockchain}.

Increasing the connectivity past the $20\%$ threshold, as shown in Figure
\ref{fig:unl-clique-consensus}, brings both cliques into agreement. However
what happens to the blockchain? The Ripple Protocol Consensus Algorithm paper
is unclear on this point. For instance, the loss and restoration of
connectivity may have been associated with network latency. The paper does
state that "nodes who’s latency grows larger than a preset bound $b$ are
removed from all UNLs"\cite[3.4.1]{ripple-consensus-paper} but does not say if
this removal is meant to be permanent\footnote{In the existing implementation
restarting the node does reload the UNL set from the config file.} nor if by
``all'' they refer to removal from UNLs on all nodes in the network.
Additionally while the author did not have sufficient time to investigate this
point fully, the source code doesn't appear to have any provision for
reorganizations of the ledger when a longer fork is detected.

Regardless, it is certain that following a fork caused by disjoint UNL cliques
the losing side simply has to discard some or all history, as seen in Figure
\ref{fig:unl-clique-post-fork}, possibly resulting in double-spends and
associated financial losses.

\begin{figure}
    \centering
    \includegraphics{figures/unl-clique-disjoint.eps}
    \caption{Disjoint UNL}
    \label{fig:unl-clique-disjoint}
\end{figure}

\begin{figure}
    \centering
    \includegraphics{figures/unl-clique-disjoint-blockchain.eps}
    \caption{Forked blockchain}
    \label{fig:unl-clique-disjoint-blockchain}
\end{figure}

\begin{figure}
    \centering
    \includegraphics{figures/unl-clique-consensus.eps}
    \caption{UNL cliques in consensus}
    \label{fig:unl-clique-consensus}
\end{figure}

\begin{figure}
    \centering
    \includegraphics{figures/unl-clique-reorganized-blockchain.eps}
    \caption{Post-fork reorganization}
    \label{fig:unl-clique-post-fork}
\end{figure}

Similarly for a UNL of $n$ nodes it would
take $(4n + 1)/5$ Byzantine failures for an incorrect transaction to be
confirmed.

Claims from ripple: consensus protocol\cite{ripple-consensus-paper} will maintain correctness so long as $f
\le (n-1)/5$ where $f$ is the number of Byzantian failures. However default UNL
list is just $5$ entries, thus $f \le (5-1)/5 = 0.8$ - no failures can be
tolerated.

@@

can talk about how cost to do a simulation attack is PoW targets; talk about slasher rule
